---
title: "Tutorial 04 - Basics of Network Visualization"
date: "CRJ 507 Social Network Analysis"
output: 
  html_document:
    df_print: paged
    theme: united
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

<style>
body {
text-align: left}
</style>

```{r, echo=FALSE}

# set the defaults for the codechunks
knitr::opts_chunk$set( eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE )

```


----

<br>

Let's be honest, network analysis is awesome. Where does that awesomeness come from? One sure reason is visualization! There is nothing like an beautiful network visualization that conveys lots of information and is aesthetically pleasing. Right? In this tutorial, you will be introduced to the basics of visualizing networks using the `gplot()` function in the `sna` package. We will also look at some approaches to building plots.

Why are you learning this? I want you to be able to render really cool networks, obviously. But, I also want you to learn how to embrace this visualization tool to convey information to audiences that might not be able to digest a table or some other data visualization medium. Let's get to work yall!

A note on the tutorials. Each tutorial in this course will contain R code which is in the “code chunks” you see below. There is also regular text. The R code chunks can be copied and pasted directly into R. As you work through this tutorial, follow along in R by coping and pasting the R code and seeing it work on your end.

<br>

----

### **Network Visualization**    

One of the great features of working with network data is the ability to ***see*** the data through visualization. Visualizing the structure of a network is helpful for discerning patterns that might be of interest. 

Douglas Luke's (2015: 47) [*A User's Guide to Network Analysis*](https://link.springer.com/book/10.1007/978-3-319-23883-8) provides several guidelines, or aesthetic principles, for what makes a graphical layout of a network easy to understand and interpret. These are:  

  + Minimize edge crossings  

  + Maximize the symmetry of the layout of nodes  

  + Minimize the variability of the edge lengths  

  + Maximize the angle between edges when they cross or join nodes  

  + Minimize the total space used for the network display  

Think about each of these suggestions. Why do they aid in visualizing the network? How do they assist in not conveying information that is not really there? Ponder these questions for a bit...

<br>

Now that you have a sense of what a good visualization should try to do, let's look at the example Luke uses as an illustration by working with the `gplot()` function in the `sna` package.  

First, we need to install the `sna` package using `install.packages( "sna" )` and load the library using `library( sna )`. Remember, if you have already installed a package then you do not need to use the `install.packages()` function. But, if you have not installed the package in a while, you should use `update.packages()` to incorporate any changes that have been made to the page.

Next, let's get the `UserNetR` package from [Douglas Luke's GitHub page](https://github.com/DougLuke/UserNetR). Since this package is not on the CRAN package repository, we need to install it from Github. We do this in four steps:  

  + First, install the devtools package using `install.packages( "devtools" )`
  
  + And then load the library for the package with `library( devtools )`
  
  + Now, install the `UserNetR` package `install_github( "DougLuke/UserNetR" )`
  
  + Finally, load the library `library( UserNetR )`  

Altogether, that looks something like this:

```{r,echo=TRUE, eval=FALSE}

# install the packages
install.packages( "sna" )
install.packages( "devtools" )

# call the libraries
library( sna )
library( devtools )

# install from Github
install_github( "DougLuke/UserNetR" )

# call that library
library( UserNetR )

```

<br>

```{r,echo=FALSE, eval=TRUE, message=FALSE}

library( sna )
library( devtools )
library( UserNetR )

```

<br>

Now, let's take a look at the `Moreno` network (see `help( Moreno, package = UserNetR )`. These data are contained in a sociogram constructed by [Jacob Moreno](https://en.wikipedia.org/wiki/Jacob_L._Moreno), and published in the New York Times in 1933 (see Moreno, J. L. 1934. *Who shall survive? A new approach to the problem of human interrelations*. Nervous and mental disease monograph series, no. 58. Washington, DC: Nervous and Mental Disease Publishing Co. for a more extensive discussion).  

```{r}

# Note that we ask the summary() function to not print out the adjacency matrix.
summary( Moreno, print.adj = FALSE )

# Now, let's compare two different plot layouts.

# Here is a circle.
gplot( Moreno, mode = "circle" )

# Here is a separate layout.
gplot( Moreno, mode = "fruchtermanreingold" )

```

<br>

Think back to the aesthetic elements we discussed above. How do these two plots differ in how well they convey the **same** information? Which one is better? Why is it better?

<br>

To think about these questions, let's plot the two layouts together. To do this we will use the `par()` function. This allows us to partition the plotting region.

```{r}

# First, we define the plot layout window.
op <- par( 
  mar = rep( 2, 4 ), # set the margins
  mfrow = c( 1, 2 )  # set the dimensions
  )

# plot the circle layout (add the main() argument for a title)
gplot( Moreno, mode = "circle", main = "Plotted as a circle" )

# plot the Fruchterman Reingold layout
gplot( Moreno, mode = "fruchtermanreingold", main = "Plotted using a spring algorithm" )

```

```{r,echo=FALSE,eval=TRUE,message=FALSE}

par( op )

```  

<br>

Again, think about our questions: How do these two plots differ in how well they convey the **same** information? Which one is better? Why is it better?

<br>

If we add information about gender, we can see a bit more that is revealed by the spring algorithm.

To do so, we use the `vertex.col=` argument with the vertex attribute *gender*. To do so, we need to access the vertex attribute. We can do this with:  
  + the `get.vertex.attribute()` function in the `sna` package. Use `library( sna )` and specify the attribute we want, like `get.vertex.attribute( Moreno,  "gender" )`
  
  + or use the shorthand for this with `network object %v% atttribute`, like `Moreno %v% "gender"`

```{r}

op <- par( mar = rep( 2, 4 ), mfrow = c( 1, 2 ) )

gplot( Moreno,
      mode = "circle",
      main = "Plotted as a circle",
      vertex.col = get.vertex.attribute( Moreno, "gender" ) # use the vertex attribute.
      )

gplot( Moreno,
       mode = "fruchtermanreingold",
       main = "Plotted using a spring algorithm",
       vertex.col = Moreno %v% "gender" # note the difference here compared to above.
       )

# let's add a legend to the plot
legend(
  "bottomleft",
  legend = c( "Male","Female" ),
  col = c( "red","black" ),
  title = "legend",
  pt.cex = 0.75,
  bty = "n",
  pch = 19 
  )

```

```{r,echo=FALSE,eval=TRUE,message=FALSE}

par( op )

```  

<br>

What is the primary story that the plot tells? Does a particular layout help us see that better?

<br>

Ok, that was a lot. Let's work our way back through the mechanics of building a plot to better get a sense of what we are doing.

<br>

### **Working with the `gplot()` function**  
Let's take a look at some of the visualization capabilities of `gplot()`. Let's start by looking at the function's help page: `?gplot`.

To see the various functionality of the function, let's work with the example of an undirected network from [Tutorial 3 - Introduction to Networks in R](https://watts-college.github.io/crj-507-spring-2024/tutorials/tutorial-03-network-data.html). 


```{r}

# define the path where the data are
url <- "https://raw.githubusercontent.com/jacobtnyoung/sna-textbook/main/data/data-undirected-example.csv"

# define the object
mat.u <- as.matrix(
  read.csv( 
    url,
    as.is = TRUE,
    header = TRUE,
    row.names = 1
    )
  )

# now, create the network
net.u <- network(
  mat.u,
  directed = FALSE
)

# take a look at the network
summary( net.u )

#NOTE: we could have wrapped all of this into two statements
net.u <- as.network( 
  as.matrix( 
    read.csv( 
      "https://raw.githubusercontent.com/jacobtnyoung/sna-textbook/main/data/data-undirected-example.csv", 
      as.is = TRUE, 
      header = TRUE, 
      row.names = 1 
      ) 
    ),
  directed = FALSE 
  )

summary( net.u )

```

<br>

Now that we have the object built, we can plot it.

```{r}

gplot( net.u )

```

<br>

Hold on, this plot shows arrows. But this network is undirected. What gives?  

That is because `gplot()` assumes a directed network. We can see this in the help menu, `?gplot`, where is shows that for the type of network, the `gmode=` argument defaults to a directed graph. To fix this we can either:  

  + manually turn off the display of arrows using the `usearrows=` argument, `gplot( net.u, usearrows = FALSE )`  

  + or indicate that the object to be plotted is a undirected graph or graph, `gplot( net.u, gmode = "graph" )`  


The `gplot()` function has a number of arguments that can be used to work try and better display the information contained in the network.

For example, we can add labels to the vertices using the `network.vertex.names()` function.

```{r}

gplot(
  net.u, 
  gmode = "graph", 
  label = network.vertex.names( net.u )
  )

```

<br>

Alternatively, we could add in a string of names for the label:  

  +  `gplot( net.u, gmode = "graph",  label = c( "Jen", "Tom", "Bob", "Leaf", "Jim" ) )`
  
<br>

Or we could read them in as an object:  

  +  `names <- c( "Jen", "Tom", "Bob", "Leaf", "Jim" )` 
  
  +  add to the plot using `gplot( net.u, gmode = "graph", label = names )`  

<br>

A great feature of R is that we can tune the graphing parameters. Here are several examples: 

  + Labels:
  
    + Add boxes around the labels, `boxed.labels = TRUE`

    + Change label size using `label.cex`, such as `label.cex = 1.5`
  
    + Color the labels using `label.col=`, such as: `label.col = "blue"`

  + Colors:
  
    + different colored names, combine `label.col=` with the `c()` function. Such as:  `label.col = c( "red", "blue", "green", "orange", "grey" )`
    
    + different colored nodes, `vertex.col=` argument. Such as: `vertex.col=c("red","blue","green","orange","grey")`
    
    + different colored edges, using `edge.col=`, such as: `edge.col=c("red","blue","green","orange","grey")`

<br>

There is a LOT of functionality to the `gplot()` function. See the arguements in the help file: `?gplot`. I would encourage you to take some time to look through it and play around with the various features.  

<br>

### **Adjusting Plot Layout**  
When a layout is generated, the results can be saved for later reuse using the `coord=` argument.

```{r}

# Set the seed for the random number generator 
# so we can always get the same plot layout.
set.seed( 507 ) 

# Define an object that will be the coordinates we want to use.
coords <- gplot( 
  net.u, 
  gmode = "graph",
  label = network.vertex.names( net.u )
  )

# Show the vertex coordinates.
coords

# Saved layouts can be used via the coord= argument:
gplot(
  net.u, 
  gmode = "graph",
  label = network.vertex.names( net.u ),
  coord = coords 
  )

```  

<br>

Cool but, *why do this?* The placement of the nodes shift when we call the `gplot()` function just due to the operation of the algorithm. Controlling where nodes are plotted in the 2-dimensional space is useful if we want to show different aspects of the plot. Note that we can have different layouts of the nodes. If we like a particular one, we can save the coordinates. 

*But*, suppose the default settings are insufficient and we want to make a few small changes. The `interactive=` argument allows for tweaking.

```{r, eval=FALSE}

# First, set up the coordinates you want.
coords <- gplot(
  net.u, 
  gmode = "graph",
  label = network.vertex.names( net.u ),
  coord = coords, 
  interactive=TRUE 
  )

```

<br>

When this renders on your system, move a few of the nodes around. Then, after you close the window it will save the coordinates.

```{r, eval=FALSE}

# Then, use these in the plot.
gplot( 
  net.u, 
  coord = coords, 
  gmode = "graph",
  label=network.vertex.names( net.u )
  ) 

```

<br>

### **Plotting the Power/Influence Network from the Prison Inmate Networks Study (PINS)**  

The [*Prison Inmate Networks Study (PINS)*](https://cjrc.la.psu.edu/research/) examines the social networks of prison inmates in a state correctional institution. The study was unique in that it was the first in nearly a century to collection sociometric data in a prison. The researchers collected data on several types of networks. 

Let's plot the *power and influence network*, which was created by asking individuals whom they believed was "powerful and influential" on the unit. We will continue working with the `gplot()` function.

We are going to do this in a few steps:

  + First, load the adjacency matrix, `data-PINS-power-w1-adj.csv`, and create an object of class `network`.
  
  + Second, load the file with Age and Race attributes, `data-PINS-w1-age-race-attributes.csv`, and assign each attribute to the network object.
  

```{r}

# define the adjacency matrix
PI.mat <- as.matrix(
  read.csv( 
    "https://raw.githubusercontent.com/jacobtnyoung/sna-textbook/main/data/data-PINS-power-w1-adj.csv",
    as.is = TRUE,
    header = TRUE,
    row.names = 1
    )
  )

# create an object of class network
PI.net <- network( PI.mat, directed = TRUE )


# define the attributes object
PI.attrs <- read.csv( 
    "https://raw.githubusercontent.com/jacobtnyoung/sna-textbook/main/data/data-PINS-w1-age-race-attributes.csv",
    as.is = TRUE,
    header = TRUE 
    )

# assign the attributes to the network
PI.net %v% "Age" <- PI.attrs[,1]
PI.net %v% "Race" <- PI.attrs[,2]

```  

<br>

Note that we used a shorthand notation: `%v%`. This is an assignment operation that tells R to assign something to the network. Specifically, `%v%` indicates the assignment to a vertex, hence the *v*.  

The operate *also* let's us pull a specific attribute. We can look at the various vertex data by using the shorthand `network %v% "attribute"`. For example:  

  + `PI.net %v% "Age"` shows the age variable.

  + `PI.net %v% "Race"` shows the race variable.

```{r}

# look at the values for age
PI.net %v% "Age" <- PI.attrs[,1]

# look at the values for race
PI.net %v% "Race" <- PI.attrs[,2]

```  

<br>

Note that we can also reference edges (i.e. `%e%`) if a network has an assigned edge. For example, we could pull the information network and assign that to the power influence network:

```{r}

# define the adjacency matrix
INFO.mat <- as.matrix(
  read.csv( 
    "https://raw.githubusercontent.com/jacobtnyoung/sna-textbook/main/data/data-PINS-info-w1-adj.csv",
    as.is = TRUE,
    header = TRUE,
    row.names = 1
    )
  )

# assign the matrix as an edge attribute
PI.net %e% "info" <- INFO.mat

```  

<br>

Think about what we did on the last line. For the power influence edges, we assigned `INFO.mat` as an attribute. This represents whether a power/influence tie was ***also*** an information network tie.

Now, we can use that information in our plot. For example: 

```{r}

gplot( PI.net,
       arrowhead.cex=0.5,
       vertex.cex = PI.net %v% "Age" )

```

<br>

YIKES!!! What is wrong?

The problem is that we need to rescale the vertex attribute so that the nodes are not too big. Let's build a function to do that and then execute the `gplot()` function:

```{r}

rescale <- function( nchar, low, high ){
  min_d <- min( nchar )
  max_d <- max( nchar )
  rscl  <- ( ( high - low )*( nchar - min_d ) ) / ( max_d - min_d ) + low
  rscl
}

```

<br>

Now, use the function we created to rescale the vertex attribute:

```{r}

gplot( PI.net, 
       arrowhead.cex=0.5,
       vertex.cex = rescale( PI.net %v% "Age", 0.5, 1.5 ) )

```

<br>

That looks better. Let's drop the isolates to help with the visualization. 

```{r}

gplot( PI.net,
       displayisolates = FALSE,
       arrowhead.cex=0.5,
       vertex.cex = rescale( PI.net %v% "Age", 0.5, 1.5 ) )

```

<br>

How about we color the edges based on whether there was a information network tie.

```{r}

gplot( PI.net,
       edge.col = PI.net %e% "info" + 1,
       displayisolates = FALSE,
       arrowhead.cex=0.5,
       vertex.cex = rescale( PI.net %v% "Age", 0.5, 1.5 ) )

```

<br>

As we build layers, we can get a fairly useful graphic that tells us a lot of information:

```{r}

gplot( PI.net,
       main="PINS Power/Influence Network", # add a title
       vertex.col = PI.net %e% "Race", # color the nodes by the Race variable
       edge.col = PI.net %e% "info" + 1, # color the edges by the information network attribute
       displayisolates = FALSE, # don't display the isolated cases
       arrowhead.cex=0.5, # augment the size of the arrowheads
       vertex.cex = rescale( PI.net %v% "Age", 0.5, 1.5 ), # size the nodes by the Age variable
       # add a subtitle
       sub="Nodes colored by Race, \n edges colored by Info net \n nodes sized by Age"
       )

```

<br>

----

### Wrapping up...

There is nothing like an beautiful network visualization that conveys lots of information and is aesthetically pleasing. Now you know how to get there! As you have seen, there is a ton of flexibility. I would encourage you to spend some time tinkering with the various arguments for the `gplot()` function to get a feel for what it can do.

You are ready to start working with your own data. If you feel comfortable with working with these data structures and plotting them, then head over and start [Lab 02 - Network Visualization](https://watts-college.github.io/crj-507-spring-2024/labs/lab-02-instructions.html). If not, that is fine. We covered a lot! Start by reviewing [Tutorial 03 - Introduction to Networks in R](https://watts-college.github.io/crj-507-spring-2024/tutorials/tutorial-03-network-data.html) as well as this tutorial again. You will be on your way to cranking out sweet plots in no time!

<br>

____

<br>

##### ***Please*** report any needed corrections to the [Issues](https://github.com/Watts-College/crj-507-spring-2024/issues/new) page. Thanks!

<br>
<br>