---
title: "Lab 05 - Bipartite Graphs and Two-Mode Networks - SOLUTIONS"
date: "CRJ 507 Social Network Analysis"
output: 
  html_document:
    df_print: paged
    theme: paper
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

<style>
body {
text-align: left}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE, eval = TRUE )
```

# Instructions

The purpose of this lab is to familiarize yourself with bipartite graphs and two-mode networks in *R*. Please review the [Bipartite Graphs/Two-Mode Networks](https://jacobtnyoung.github.io/sna-textbook/ch-10-bipartite-graphs.html) chapter of the [SNA Textbook](https://jacobtnyoung.github.io/sna-textbook/index.html) as well as the [ipartite Graphs and Two-Mode Networks in R]() tutorial prior to beginning this lab.

<br>

## Part I: A Conspiracy Network

For this part of the lab, you will use data collected from [Paul Revere's Ride](https://global.oup.com/ushe/product/paul-reveres-ride-9780195098310?cc=us&lang=en&), by David Fischer. In the book, Fischer documents Reveres connections through various affiliations in locations and how these influenced history. The Paul Revere conspiracy dataset concerns relationships between 254 people and their affiliations with seven different organizations in Boston. The dataset refers to Paul Revere, who was responsible for organizing a local militia of Boston's revolutionary movement. The network is two-mode, with 254 actors and 7 organizations ("events"). We will refer to this network as the `PaulRevereNet`.

<br>

For the `PaulRevereNet` network, do the following:  

1. Plot the network using the `gplot()` function.  
2. Calculate the density of the network.
3. Interpret the density of the network.
4. Calculate the degree centrality scores for each set of nodes.  
5. Calculate the mean degree centrality score.  
6. Calculate the standardized degree centrality scores for each set of nodes.  
7. Calculate the mean standardized degree centrality score.  
8. Plot the network using the standardized degree centrality scores.
9. Using the vertex attribute `individual.names`, discuss which individual is most central in the network.
10. Using the vertex attribute `place.names`, discuss which location is most central in the network.

<br>

## Part II: A Kidnapping Network

For this part of the lab, you will use the Philippine Kidnappings Data which is a collection of relationships on the Abu Sayyaf Group (ASG), a violent non-state actor operating in the Southern Philippines. In particular, this data is related to the Salast movement that has been founded by Aburajak Janjalani, a native terrorist of the Southern Philippines in 1991. ASG is active in kidnapping and other kinds of terrorist attacks. The object is a two-mode network where actors are terrorist kidnappers and the terrorist events they have attended. We will refer to this as the `PhilKidnapNet` network.

For this part of the lab, use the `PhilKidnapNet` to repeat steps 1-8 in Part I.

<br>
<br>

----


# **Solutions**  

<br>

## Setup

```{r, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

```

## Part I

<br>

First, let's read in the `PaulRevereNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-paul-revere-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PaulRevereNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# load the libraries we need
library( sna )
library( network )

# define the path location for the file
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-paul-revere-net.rds"
PaulRevereNet <- readRDS( url( loc ) )

# look at the network
PaulRevereNet

```

<br>

#### 1. Plot the network using the `gplot()` function.  

<br>

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,             # our network to plot
  gmode = "twomode",         # indicate it is two modes
  usearrows = FALSE,         # turn off the arrowheads
  vertex.cex=2,              # size the nodes     
  main="Paul Revere Network" # add a title
)

```

<br>

#### 2. Calculate the density of the network.

<br>

```{r}

# create a matrix from the network object
PaulRevereMat <- as.matrix( PaulRevereNet )

# identify the number of edges in the graph
L <- sum( PaulRevereMat )

# identify the number of actors in the example
N <- dim( PaulRevereMat )[1]

# identify the number of events in the example
M <- dim( PaulRevereMat )[2]

# calculate the density
density.PaulRevere <- L / ( N * M )

# check it out
density.PaulRevere

```

<br>

#### 3. Interpret the density of the network.

The density of for the Paul Revere network indicates that `r round( density.PaulRevere, 2 ) * 100` percent of the ties that could be observed, were observed. This also means that `r round( density.PaulRevere, 2 ) * 100` of the ties that could have occurred were not observed. Overall, we can say that the network is fairly *sparse*.

<br>

#### 4. Calculate the degree centrality scores for each set of nodes.  

```{r}

# raw scores for actors
actor.deg <- rowSums( PaulRevereMat )
actor.deg

# raw scores for events
event.deg <- colSums( PaulRevereMat )
event.deg

```

<br>

#### 5. Calculate the mean degree centrality scores and interpret the values.  

```{r}

# mean degree for actors
mean.actor.deg <- L / N

# mean degree for events
mean.event.deg <- L / M

# an alternative is to just use the mean() function with the degree data
mean( actor.deg )
mean( event.deg )

```

<br>

For the first set of nodes, the individuals who could attend events at locations, the mean degree is `r mean( actor.deg )`, meaning that on average, individuals attended `r round( mean( actor.deg ), 2 )` events.  

For the second set of nodes, the locations where individuals could meet, the mean degree is `r mean( event.deg )`, meaning that each location, on average, had `r round( mean( event.deg ), 2 )` attendees.  

<br>

#### 6. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# standardized score for actors
actor.deg / M

# standardized score for events
event.deg / N

```

<br>

#### 7. Calculate the mean standardized degree centrality score.  

```{r}

# mean standardized score for actors
mean( actor.deg / M )

# mean standardized score for events
mean( event.deg / N )

```

<br>

Notice that these are the same because we adjusted for the different sizes of *M* and *N* in the network.

<br>

#### 8. Plot the network using the standardized degree centrality scores.

```{r}

# define the standardized scores for actors
actor.size <- actor.deg / M

# define the standardized scores for events
event.size <- event.deg / N

# combine these to use in the plot
v.size <- c( actor.size, event.size )

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,              # our network to plot
  gmode = "twomode",          # indicate it is two modes
  usearrows = FALSE,          # turn off the arrowheads
  main="Paul Revere Network", # add a title
  vertex.cex = v.size + 0.5   # set the size (add 0.5 so it is not too small) 
)

```

<br>

#### 9. Using the vertex attribute `people.names`, discuss which individual is most central in the network.

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,                               
  gmode = "twomode",                           
  usearrows = FALSE,                           
  main="Paul Revere Network",                  
  vertex.cex = v.size + 0.5,                   
  label = PaulRevereNet %v% "people.names",    # add the individual names
  label.cex = 0.4,                             # change the label size
  label.pos = 5                                # set the label position
)

```

<br>

This is hard to see! One way around this is to just show labels for the those nodes that have degree centrality scores at or above the mean. Let's see how we can do this.

First, identify who is the most central person. We can do this using the `sort()` function.

```{r}

sort( actor.deg / M )

```

We see it is Paul Revere!

Now, we want to change the names to only show Paul Revere's name.

```{r}

# create a vector of names to change
actor.adj.names <- PaulRevereNet %v% "people.names"

# if the name is not Paul Revere, make it NA
actor.adj.names[ actor.adj.names != "Paul Revere" ] <- NA

```

Now that we have adjusted the names, we can plot it.

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,              
  gmode = "twomode",          
  usearrows = FALSE,          
  main="Paul Revere Network", 
  vertex.cex = v.size + 0.5,  
  label = actor.adj.names,    # here we use the new names
  label.cex = 0.8,            # change the label size
  label.pos = 5               
)

```

<br>

#### 10. Using the vertex attribute `place.names`, discuss which location is most central in the network.

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,                            
  gmode = "twomode",                        
  usearrows = FALSE,                        
  main="Paul Revere Network",               
  vertex.cex = v.size + 0.5,                
  label = PaulRevereNet %v% "place.names",  # use the place names
  label.cex = 1.2,                          
  label.pos = 5                             
)

```

Again, this is a bit tricky to see from the plot. *But*, we can examine the degree centrality scores as we did above with the `sort()` function.

```{r}

# sort the degree centrality scores in descending order
sort( event.deg / N )

```

We see it is the "TeaParty" location (although the "LondonEnemies" location is very close in size). We can see this a bit better in our plot by adjusting the color of the edges **and** expanding the size of the labels using the `rescale()` function.

```{r}

# create the function
rescale <- function( nchar, low, high ){
  min_d <- min( nchar )
  max_d <- max( nchar )
  rscl  <- ( ( high - low )*( nchar - min_d ) ) / ( max_d - min_d ) + low
  rscl
}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PaulRevereNet,                            
  gmode = "twomode",                        
  usearrows = FALSE,                        
  main="Paul Revere Network",               
  vertex.cex = v.size + 0.5,                
  label = PaulRevereNet %v% "place.names",  
  label.cex = rescale( v.size, 0.3, 1.8 ), # pass the rescale function to the label.cex argument                          
  label.pos = 5,                            
  edge.col = "grey80"                      # change the color of the edges
)

```

<br>

## Part II

<br>

First, let's read in the `PhilKidnapNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-philippine-kidnappings-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PhilKidnapNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# clear the workspace so we don't use any duplicate names
rm( list = ls() )

# define the path location for the file
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-philippine-kidnappings-net.rds"
PhilKidnapNet <- readRDS( url( loc ) )

# look at the network
PhilKidnapNet

```

<br>

#### 1. Plot the network using the `gplot()` function.  

<br>

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PhilKidnapNet,             
  gmode = "twomode",         
  usearrows = FALSE,         
  vertex.cex = 1.2,              
  main="Philippe Kidnapping Network"
)

```

<br>

#### 2. Calculate the density of the network.

<br>

```{r}

# create a matrix from the network object
PhilKidnapMat <- as.matrix( PhilKidnapNet )

# identify the number of edges in the graph
L <- sum( PhilKidnapMat )

# identify the number of actors in the example
N <- dim( PhilKidnapMat )[1]

# identify the number of events in the example
M <- dim( PhilKidnapMat )[2]

# calculate the density
density.PhilKidnap <- L / ( N * M )

# check it out
density.PhilKidnap

```

<br>

#### 3. Interpret the density of the network.

The density of for the Paul Revere network indicates that `r round( density.PhilKidnap, 3 ) * 100` percent of the ties that could be observed, were observed. This also means that `r round( density.PhilKidnap, 3 ) * 100` of the ties that could have occurred were not observed. Overall, we can say that the network is very *sparse*.

<br>

#### 4. Calculate the degree centrality scores for each set of nodes.  

```{r}

# raw scores for actors
actor.deg <- rowSums( PhilKidnapMat )
actor.deg

# raw scores for events
event.deg <- colSums( PhilKidnapMat )
event.deg

```

<br>

#### 5. Calculate the mean degree centrality scores and interpret the values.  

```{r}

# mean degree for actors
mean.actor.deg <- L / N

# mean degree for events
mean.event.deg <- L / M

# an alternative is to just use the mean() function with the degree data
mean( actor.deg )
mean( event.deg )

```

<br>

For the first set of nodes, the individuals who could attend events at locations, the mean degree is `r mean( actor.deg )`, meaning that on average, individuals attended `r round( mean( actor.deg ), 2 )` events.  

For the second set of nodes, the locations where individuals could meet, the mean degree is `r mean( event.deg )`, meaning that each location, on average, had `r round( mean( event.deg ), 2 )` attendees.  

<br>

#### 6. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# standardized score for actors
actor.deg / M

# standardized score for events
event.deg / N

```

<br>

#### 7. Calculate the mean standardized degree centrality score.  

```{r}

# mean standardized score for actors
mean( actor.deg / M )

# mean standardized score for events
mean( event.deg / N )

```

<br>

Notice that these are the same because we adjusted for the different sizes of *M* and *N* in the network.

<br>

#### 8. Plot the network using the standardized degree centrality scores.

First, let's reload the `rescale()` function to help with the plot sizes.

```{r}

# create the function
rescale <- function( nchar, low, high ){
  min_d <- min( nchar )
  max_d <- max( nchar )
  rscl  <- ( ( high - low )*( nchar - min_d ) ) / ( max_d - min_d ) + low
  rscl
}

```

Now, we can create the vector of sizes and make the plot.

```{r}

# define the standardized scores for actors
actor.size <- actor.deg / M

# define the standardized scores for events
event.size <- event.deg / N

# combine these to use in the plot
v.size <- c( actor.size, event.size )

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PhilKidnapNet,              
  gmode = "twomode",          
  usearrows = FALSE,          
  main="Philippe Kidnapping Network",
  vertex.cex = rescale( v.size, 0.5, 1.8 )   
)

```

<br>

#### 9 & 10. Using the vertex attribute `vertex.names`, discuss which individual and which event are the most central in the network.

```{r}

# set the seed to reproduce the plot layout
set.seed( 507 )

# execute the plot
gplot(
  PhilKidnapNet,                               
  gmode = "twomode",                           
  usearrows = FALSE,                           
  main="Philippe Kidnapping Network",
  vertex.cex = rescale( v.size, 0.5, 1.8 ),   
  label = PhilKidnapNet %v% "vertex.names", # add the individual names
  label.cex = rescale( v.size, 0.3, 1.8 ),  # pass the rescale function to the label.cex argument
  label.pos = 5,                            # set the label position
  edge.col = "grey80"                       # change the color of the edges
)

```

From the plot, we can see that `r which( actor.deg / M == max( actor.deg / M ) )` is the most central actor indicating that this person attended the most meetings. And, we see that `r which( event.deg / N == max( event.deg / N ) )` is the most central event meaning that this meeting had the most attendees.

<br>
<br>

----

##### ***Please*** report any needed corrections to the [Issues](https://github.com/Watts-College/crj-507-template/issues) page. Thanks!

<br>
<br>

###### ***Last updated `r format(Sys.time(), '%d %B, %Y')`***