---
title: "Final Project"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( echo = FALSE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

```{r, eval=TRUE, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

# load the libraries you will need
library( network )
library( sna )


# Load the networks

# set the url
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-PHX-arrest-2023-net.rds"

# for the 2023
PhxArrestNet2023 <- readRDS( url( loc ) )

# set the url
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-PHX-arrest-2022-net.rds"

# for the 2022
PhxArrestNet2022 <- readRDS( url( loc ) )

```


## Part I: Two-Mode Network

*What can you tell me about this network?* 

YOUR ANSWER TO THIS QUESTION HERE (THEN DELETE THIS LINE)

```{r, echo=FALSE}

# your code here for creating objects here

```

<br>

## Part II: One-Mode Network

*Who are the central actors in this network?* 

YOUR ANSWER TO THIS QUESTION HERE (THEN DELETE THIS LINE)

```{r, echo=FALSE}

# your code here for creating objects here

```

<br>

## Part III: Comparison

*How has the co-offending network changed from 2022 to 2023?* 

YOUR ANSWER TO THIS QUESTION HERE (THEN DELETE THIS LINE)

```{r, echo=FALSE}

# your code here for creating objects here

```

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
