---
title: "Lab 06 - Projection of Bipartite Graphs"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

# Getting Started

<br>

## Setup

```{r, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

```

## Part I

<br>

First, let's read in the `PaulRevereNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-paul-revere-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PaulRevereNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# Your code here

```

<br>

#### 1. Create an object of class `matrix` from the `PaulRevereNet` network.
  
<br>

```{r}

# Your code here

```

<br>

#### 2. Create the unipartite graph that represents the "person" matrix.

<br>

```{r}

# Your code here

```

<br>

#### 3. Create the unipartite graph that represents the "group" matrix.

<br>

```{r}

# Your code here

```

<br>

#### 4. Plot all three graphs.

Let's use the `edge.rescale()` function to help us here. This function returns a weighted edgelist that can be used to aid with plotting.

```{r}

edge.rescale <- function( uniMat, low, high ){
  diag( uniMat ) <- 0
  min_w <- min( uniMat[uniMat != 0] )
  max_w <- max( uniMat[uniMat != 0] )
  rscl <- ( ( high-low )  * ( uniMat[uniMat != 0] - min_w ) ) / ( max_w - min_w ) + low
  rscl
}

edge.shade <- function( uniMat ){
  net.edges <- edge.rescale( uniMat, 0.01, 1 )
  vec.to.color <- as.vector( abs( net.edges ) )
  vec.to.color <- 1 - vec.to.color # subtract 1 to flip the grey function scale.
  edge.cols <- grey( vec.to.color )
  return( edge.cols )
}


```

Now, plug it into the plot.

```{r, fig.width=10, fig.height=10, fig.fullwidth=TRUE}

# set the plot regions to ease with visualization
par( 
  mfrow = c( 2, 2 ),
  mar = c( 2, 1, 4, 1)
  )

# set the seed to reproduce the plot
set.seed( 507 )


# Your code for the plots here

```
 
<br>

## Part II

<br>

First, let's read in the `PhilKidnapNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-philippine-kidnappings-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PhilKidnapNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# Your code here

```

<br>

#### 1. Create an object of class `matrix` from the `PhilKidnapNet` network.
  
<br>

```{r}

# Your code here

```

<br>

#### 2. Create the unipartite graph that represents the "person" matrix.

<br>

```{r}

# Your code here

```

<br>

#### 3. Create the unipartite graph that represents the "group" matrix.

<br>

```{r}

# Your code here

```

<br>

#### 4. Plot all three graphs.

Let's use the `edge.rescale()` function to help us here. This function returns a weighted edgelist that can be used to aid with plotting.

```{r}

edge.rescale <- function( uniMat, low, high ){
  diag( uniMat ) <- 0
  min_w <- min( uniMat[uniMat != 0] )
  max_w <- max( uniMat[uniMat != 0] )
  rscl <- ( ( high-low )  * ( uniMat[uniMat != 0] - min_w ) ) / ( max_w - min_w ) + low
  rscl
}

edge.shade <- function( uniMat ){
  net.edges <- edge.rescale( uniMat, 0.01, 1 )
  vec.to.color <- as.vector( abs( net.edges ) )
  vec.to.color <- 1 - vec.to.color # subtract 1 to flip the grey function scale.
  edge.cols <- grey( vec.to.color )
  return( edge.cols )
}


```

Now, plug it into the plot.

```{r, fig.width=10, fig.height=10, fig.fullwidth=TRUE}

# set the plot regions to ease with visualization
par( 
  mfrow = c( 2, 2 ),
  mar = c( 2, 1, 4, 1)
  )

# set the seed to reproduce the plot
set.seed( 507 )


# Your code for the plots here

```

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
