---
title: "Lab 02 - Network Visualization"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6)

```

<br>

```{r load_packages}
# make sure these are installed on your machine before knitting
# install.packages( "network" )
# install.packages( "sna" )
#
# Note that you cannot knit with an install.packages() function
# called in your RMD file. You only have to install packages
# once, then they are loaded using the library() function.

library( network )
library( sna )

```

<br>

As indicated in the instructions, for *each* network you created, do the following:  

1. Create an adjacency matrix of the network and save it as a .csv file.
2. Import the network into *R* using the `read.csv()` function or create the network using the `matrix()` function.
3. Create an object of class `network`. 
4. Plot the network using the `gplot()` function.    
5. Modify the plot using at least three arguments in the `gplot()` function.  

As a refresher, review the example below.

<br>

## Example: 

If you are importing your network, use the `read.csv()` function:

```{r import-example, echo = TRUE, eval = TRUE}

# define the path
url <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-undirected-example.csv"

# import the undirected matrix
mat.u <- read.csv( 
  url,
  as.is = TRUE,
  header = TRUE,
  row.names = 1 
  )

```

If you are creating your network, use the `matrix()` function:

<br>

```{r create-example, echo = TRUE, eval = TRUE}

# created a directed matrix
mat.d <- matrix(
  c(0,1,0,0,0,1,0,1,0,0,0,1,0,1,1,0,0,1,0,1,0,0,1,1,0),
  nrow=5,
  byrow=TRUE
  )

# attach row names
rownames( mat.d ) <- c( "Jen","Tom","Bob","Leaf","Jim" )

# attach column names
colnames( mat.d ) <- c( "Jen","Tom","Bob","Leaf","Jim" )

# print out the object
mat.d

```

<br>

Create an object of class `network` using the `as.network()` function. 

```{r coerce-example, echo =  TRUE, eval = TRUE}

# coerce the matrix object
net.d <- as.network( 
  mat.d,
  directed = TRUE # set this to false for an directed network; FALSE for an undirected network
  )

# print the object to see the properties
net.d

```


<br>

Finally, let's plot the networks using the `gplot()` function. Note that we can use either the object `mat.u` which is of class `matrix` or the object `net.u` which is of class `network`.

```{r create-undirected-plot, echo = TRUE, eval = TRUE}

# produce the plot
gplot(
  mat.u,                          # this is the network to plot
  gmode = "graph",                # let it know it is undirected    
  vertex.col = "lightblue",       # color the nodes
  vertex.cex = 1.2,               # size of the nodes (bigger number is bigger nodes)
  edge.col = "grey80",            # color the ties
  edge.lwd = 1.2,                 # the width of the ties (bigger is wider)
  main = "Plot of undirected net" # add a title
)

```

<br>

```{r create-directed-plot, echo = TRUE, eval = TRUE}

# produce the plot
gplot(
  mat.d,                          # this is the network to plot
  vertex.col = "firebrick",       # color the nodes
  vertex.cex = 1.2,               # size of the nodes (bigger number is bigger nodes)
  edge.col = "grey80",            # color the ties
  edge.lwd = 1.2,                 # the width of the ties (bigger is wider)
  main = "Plot of a directed net" # add a title
)

```

<br>

These were some examples to help you get going. Feel free to use them to aid in working through the lab.

```{r clear-workspace, echo = TRUE, eval = TRUE}

# clear the workspace before you start working on your sections

rm( list = ls() )

```

----

# Part I

<br>

## First Network

<br>

### Step 2

Import the network into *R* using the `read.csv()` function or create the network using the `matrix()` function.

```{r}

# Your code here

```

<br>

### Step 3

Create an object of class `network`. 

```{r}

# Your code here

```

<br>

### Steps 4 & 5

Create an object of class `network`. Plot the network using the `gplot()` function. Modify the plot using at least three arguments in the `gplot()` function.    

<br>

----

# Part II

<br>

## Second Network

<br>

### Step 2

Import the network into *R* using the `read.csv()` function or create the network using the `matrix()` function.

```{r}

# Your code here

```

<br>

### Step 3

Create an object of class `network`. 

```{r}

# Your code here

```

<br>

### Steps 4 & 5

Create an object of class `network`. Plot the network using the `gplot()` function. Modify the plot using at least three arguments in the `gplot()` function.    

<br>

```{r}

# Your code here

```

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
