---
title: "Lab 04 - Closeness/Betweenness Centrality"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

# Getting Started

## Setup

```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I

<br>

First, let's read in the `LondonGangNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-london-gang-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `LondonGangNet` object as one of class `network`.

<br>

```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

# load the libraries we need
library( sna )
library( network )

# define the path location for the file
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-london-gang-net.rds"
LondonGangNet <- readRDS( url(loc ) )

# look at the network
LondonGangNet

```

<br>

#### 1. Calculate the degree, closeness, and betweenness centrality scores for each actor.  

<br>

```{r}

# use the degree() function in the sna package
deg.gang <- degree( 
  LondonGangNet,
  gmode = "graph", 
  cmode = "degree"
  )

```

Now let's calculate the closeness and betweenness scores. Recall that the `closeness()` function reports the standardized closeness centrality by default. If we want the unstandardized closeness, we can just divide the result returned by `closeness()` by *g-1*.

```{r}

# set the number of nodes in the network
g <- dim( as.matrix( LondonGangNet ) )[1]

# Your code here

```

<br>

#### 2. Calculate the standardized degree, closeness, and betweenness centrality scores for each actor.  

```{r}

# Your code here

```

<br>

#### 3. Calculate the mean degree, closeness, and betweenness centrality score.

```{r}

# use the mean() function to calculate the means

# Your code here
 
```

<br>

#### 4. Compare the differences for the three mean centrality scores for each measure of centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 5. Calculate the graph centralization for degree, closeness, and betweenness centrality.  

```{r}

# Your code here

```

<br>

#### 6. Discuss the graph centralization score for closeness centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 7. Discuss the graph centralization score for betweenness centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 8. Compare the differences for the three graph centralization scores for each measure of centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 9. Plot the network three times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for degree centrality, a plot for closeness centrality, and a plot for betweenness centrality).

Be sure to use the `rescale()` function from the [betweenness centrality]() tutorial to size the nodes.

```{r}

# for sizing the nodes
rescale <- function( nchar, low, high ){
  min_d <- min( nchar )
  max_d <- max( nchar )
  rscl  <- ( ( high - low )*( nchar - min_d ) ) / ( max_d - min_d ) + low
  rscl
}

# Your code here

```

<br>

#### 10. Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plot.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

## Part II

<br>

First, let's read in the `CocaineDealingNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-cocaine-dealing-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Above, we needed to make sure the `sna` and `network` packages were loaded, but we loaded them above so we do not need to repeat the step here.

<br>

```{r, echo=TRUE, eval=TRUE}

# define the path location for the file
loc2 <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-cocaine-dealing-net.rds"
CocaineDealingNet <- readRDS( url(loc2 ) )

# look at the network
CocaineDealingNet

```

<br>

#### 1. Calculate the indegree, outdegree, closeness, and betweenness centrality scores for each actor.  
  
```{r}

# use the degree() function in the sna package.
ideg.coc <- degree( 
  CocaineDealingNet,
  gmode = "digraph", # note the change to a directed graph  
  cmode = "indegree" # ask for the indegree
  )

odeg.coc <- degree( 
  CocaineDealingNet,
  gmode = "digraph",  # note the change to a directed graph
  cmode = "outdegree" # toggle to return the outdegree
  )

```

Now let's calculate the closeness and betweenness scores. Again, recall that the `closeness()` function reports the standardized closeness centrality by default. If we want the unstandardized closeness, we can just divide the result returned by `closeness()` by *g*. Note, that since the graph is directed, we divide by *g*, not *g-1*.

```{r}

# set the number of nodes in the network
g <- dim( as.matrix( CocaineDealingNet ) )[1]

# Your code here

```

<br>

#### 2. Calculate the standardized indegree, outdegree, closeness, and betweenness centrality scores for each actor.  

```{r}

# Your code here

```

<br>

#### 3. Calculate the mean indegree, outdegree, closeness, and betweenness centrality score.

```{r}

# use the mean() function to calculate the means

# Your code here

```

<br>

#### 4. Compare the differences for the four mean centrality scores for each measure of centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 5. Calculate the graph centralization for indegree, outdegree, closeness, and betweenness centrality.   

```{r}

# Your code here

```

<br>

#### 6. Discuss the graph centralization score for closeness centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 7. Discuss the graph centralization score for betweenness centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

#### 8. Compare the differences for the three graph centralization scores for each measure of centrality.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 9. Plot the network four times using the `gplot()` function where each plot sizes the nodes by the centrality score (i.e. a plot for indegree centrality, a plot for outdegree centrality, a plot for closeness centrality, and a plot for betweenness centrality).

<br>

```{r}

# Your code here

```

<br>

#### 10. Examine the plots. Describe how the differences you noted between the centrality measures above are seen in the plot.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>
<br>

----

<br>
<br>

## Part III:

<br>

*Add your code setup here according to the instructions above, then DELETE THESE STATEMENT*

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
