---
title: "Lab 05 - Bipartite Graphs and Two-Mode Networks"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    df_print: paged
    theme: journal
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: show
    self_contained: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set( echo = TRUE,
                       eval = TRUE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6 )

```

<br>

# Getting Started

## Setup

```{r, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

# clear the workspace
rm( list = ls() )

```

<br>

## Part I

<br>

First, let's read in the `PaulRevereNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-paul-revere-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PaulRevereNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# load the libraries we need
library( sna )
library( network )

# define the path location for the file
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-paul-revere-net.rds"
PaulRevereNet <- readRDS( url( loc ) )

# look at the network
PaulRevereNet

```

<br>

#### 1. Plot the network using the `gplot()` function.  

<br>

```{r}

# Your code here

```

<br>

#### 2. Calculate the density of the network.

<br>

```{r}

# Your code here

```

<br>

#### 3. Interpret the density of the network.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 4. Calculate the degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

#### 5. Calculate the mean degree centrality scores and interpret the values.  

```{r}

# Your code here

```

<br>

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 6. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

#### 7. Calculate the mean standardized degree centrality score.  

```{r}

# Your code here

```

<br>

#### 8. Plot the network using the standardized degree centrality scores.

```{r}

# Your code here

```

<br>

#### 9. Using the vertex attribute `people.names`, discuss which individual is most central in the network.

```{r}

# Your code here

```

<br>

#### 10. Using the vertex attribute `place.names`, discuss which location is most central in the network.

```{r}

# Your code here

```

<br>
<br>

----

## Part II

<br>

First, let's read in the `PhilKidnapNet` network. This is stored as an .rds document in the data folder on the [SNA Textbook](https://github.com/jacobtnyoung/sna-textbook/tree/main/data) site. We will use the `readRDS()` function, with the [file path](https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-philippine-kidnappings-net.rds), to load the file. Since we are calling a url, we need to use the `url()` function as well.  

Finally, we need to make sure the `sna` and `network` packages are loaded, using `library()`, so that R recognizes the `PhilKidnapNet` object as one of class `network`.

<br>

```{r, message=FALSE, warning=FALSE}

# clear the workspace so we don't use any duplicate names
rm( list = ls() )

# define the path location for the file
loc <- "https://github.com/jacobtnyoung/sna-textbook/raw/main/data/data-philippine-kidnappings-net.rds"
PhilKidnapNet <- readRDS( url( loc ) )

# look at the network
PhilKidnapNet

```

<br>

#### 1. Plot the network using the `gplot()` function.  

<br>

```{r}

# Your code here

```

<br>

#### 2. Calculate the density of the network.

<br>

```{r}

# Your code here

```

<br>

#### 3. Interpret the density of the network.

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 4. Calculate the degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

#### 5. Calculate the mean degree centrality scores and interpret the values.  

```{r}

# Your code here

```

<br>

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>

#### 6. Calculate the standardized degree centrality scores for each set of nodes.  

```{r}

# Your code here

```

<br>

#### 7. Calculate the mean standardized degree centrality score.  

```{r}

# Your code here

```

<br>

#### 8. Plot the network using the standardized degree centrality scores.

```{r}

# Your code here

```

<br>

#### 9 & 10. Using the vertex attribute `vertex.names`, discuss which individual and which event are the most central in the network.

```{r}

# Your code here

```

*add your interpretation here (THEN DELETE THIS STATEMENT)*

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Upload both your `.RMD` and `.HTML` files to the appropriate link for this assignment on the Canvas page for this course.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  
